---
title: "Problem Set 1"
author: "Amanda Carpenter"
format: pdf
---

## Problem Set 1

```{r}
library(tidyverse)
mtcars_clean <- read_csv("~/Documents/GitHub/data-viz-summer-25/work/data/mtcars_clean.csv")
```


```{r}
ggplot(mtcars_clean, aes(x = factor(am))) +
  geom_bar(fill = "darkgreen", color = "gray20", linewidth = .5) +
  scale_x_discrete(labels = c("0" = "Automatic", "1" = "Manual")) +
  labs(
    title = "Frequency of Transmission Types",
    x = "",
    y = "Count"
  ) +
  theme_minimal()
```


# Bar Chart
Bar charts are useful for presenting categorical data. Bar charts report the frequencies for categorical data. 


```{r}
ggplot(mtcars_clean, aes(x = hp, y = mpg)) +
  geom_point(color = "steelblue") +
  labs(
    subtitle = "Relationship between Horsepower and Miles Per Gallon",
    x = "Horsepower",
    y = "MPG"
  ) +
  theme_minimal()
```


# Scatter Plot
Scatter plots are useful for presenting the visual relationship between two variables. Scatter plots show the association between two variables.

```{r}
ggplot(mtcars_clean, aes(x = efficiency)) +
  geom_histogram(binwidth = 2, fill = "purple", color = "black") +
  theme_minimal()
```


# Histogram
Histograms are useful for presenting continuous data. Histograms show the distribution of continuous data. They can provide insight to the normality of the distribution. 